---
title: "Nematode Functional Diversity"
author: "Wei, C.-L., Liao, J.-X."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reguired package
```{r}
library(nema)
library(FD)
library(ggplot2)
library(reshape2)
library(vegan)
library(plyr)
library(viridis)
library(doBy)
library(cowplot)
library(knitr)
library(nlme)
library(MuMIn)
```

# ggplot theme setting
```{r}
large <- theme(legend.title = element_text(size=15),
        legend.text = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text = element_text(size=15),
        strip.text = element_text(size=15))

rotate <- theme(axis.text.x = element_text(angle = 30, hjust=0.5))

no_strip <- theme(strip.background = element_rect(colour=NA, fill=NA),
                  strip.text = element_text(colour=NA))
```

# Preparing nematode and environmental data
Relative abundance of nematode species from 100 randomly slected individuals was scaled to the total nematode abundance in each sample.
```{r}
# Convert to species-by-sample matrix
b <- acast(nema_abund, Cruise+Station+Deployment+Tube+Subcore~Genus+Species, fill=0)
# Multiply the relative abundance by total abundance
b <- ceiling(decostand(b, "total")*nema_total$Abundance)

# Match environmental data to neamtode sample
e <- ldply(strsplit(row.names(b), split="_"))
e <- cbind(with(e, paste(V1, V2, sep="_")), e[,-1:-2])
names(e) <- c("Cruise", "Station", "Deployment", "Tube", "Subcore")
id1 <- with(e, paste(Cruise, Station, Deployment))
id2 <- with(nema_cruise, paste(Cruise, Station, Deployment))
e <- cbind(e, nema_cruise[match(id1, id2), -3:-5])

# Define depth zone
depth.bk <- c(200, 400, 600, 800, 1100)
depth.lab <- c("200-400", "400-600", "600-800", "800-1100")
e$Zone <- cut(e$Depth, breaks=depth.bk, labels=depth.lab)

# Match the trait data to the species-by-sample matrix
s <- nema_species[match(colnames(b), with(nema_species, paste(Genus, Species, sep="_"))),]
```

# Estimate Functional Hill Numbers
* Bootstrap confidence interval (1000 iteration)
```{r, eval=FALSE}
tr <- s[, c("Buccal", "Tail", "Life.history")]
row.names(tr) <- colnames(b)
dis <- gowdis(tr)
ab <- as.list(as.data.frame(t(b)))
Q <- raoQ(ab, dis, datatype = "abundance")
fd <- FunD(ab, dis, tau=Q, q=0:2, boot=1000, datatype = "abundance")
save(fd, file="../rda/FunHillNumbers.rda")
```

```{r}
load("../rda/FunHillNumbers.rda")
fd <- subset(fd$forq, tau=="dmean")
```


## q = 0, tau = RaoQ of functional richness
```{r, fig.width=6, fig.height=4}
ggplot(data=cbind(subset(fd, q==0), e), 
       aes(x=Depth, y=estimate, ymin=LCL, ymax=UCL, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  stat_smooth(data=subset(cbind(subset(fd, q==0), e), Habitat=="Slope"), method="lm", fill="gray60", colour=viridis(2)[2])+
  labs(x="Depth (m)", y="Functional Diversity (q = 0)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large

# Linear regression
hl <- splitBy(~Habitat, cbind(subset(fd, q==0), e))
lapply(hl, FUN=function(x)summary(lm(estimate~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = estimate ~ Habitat*Depth, random = ~1|Cruise, data=cbind(subset(fd, q==0), e), method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)

# Normalized enivronmental data to mean zero and unit variance
es <- as.data.frame(scale(e[, c(-1:-9, -11:-13, -35)]))

# General linear Models
dat <- cbind(subset(fd, q==0), es)
f <- gls(estimate ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
kable(summary(b)$tTable)
# Relative importance
kable(summary(ma)$sw)
```


```{r, fig.width=6, fig.height=4}
out <- summaryBy(estimate~Cruise+Station+Habitat+Zone, data=cbind(subset(fd, q==0), e), FUN=c(mean, sd))
out$Date <- factor(out$Cruise, labels=c("2015-08", "2015-11"))
names(out)[5:6] <- c("estimate", "sd")

ggplot(data=out, aes(x=Zone, y=estimate, ymin=estimate, ymax=estimate+sd, fill=Date))+
  geom_errorbar(position="dodge")+
  geom_bar(stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Functional Diversity (q = 0)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

## q = 1, tau = RaoQ of effective number of typical functional group
```{r, fig.width=6, fig.height=4}
ggplot(data=cbind(subset(fd, q==1), e), 
       aes(x=Depth, y=estimate, ymin=LCL, ymax=UCL, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  #stat_smooth(method="lm", fill="gray60")+
  labs(x="Depth (m)", y="Functional Diversity (q = 1)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large

# Linear regression
hl <- splitBy(~Habitat, cbind(subset(fd, q==1), e))
lapply(hl, FUN=function(x)summary(lm(estimate~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = estimate ~ Habitat*Depth, random = ~1|Cruise, data=cbind(subset(fd, q==1), e), method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)

# General linear Models
dat <- cbind(subset(fd, q==1), es)
f <- gls(estimate ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
kable(summary(b)$tTable)
# Relative importance
kable(summary(ma)$sw)
```

```{r, fig.width=6, fig.height=4}
out <- summaryBy(estimate~Cruise+Station+Habitat+Zone, data=cbind(subset(fd, q==1), e), FUN=c(mean, sd))
out$Date <- factor(out$Cruise, labels=c("2015-08", "2015-11"))
names(out)[5:6] <- c("estimate", "sd")

ggplot(data=out, aes(x=Zone, y=estimate, ymin=estimate, ymax=estimate+sd, fill=Date))+
  geom_errorbar(position="dodge")+
  geom_bar(stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Functional Diversity (q = 1)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

## q = 2, tau = RaoQ of effective number of dominant functional group
```{r, fig.width=6, fig.height=4}
ggplot(data=cbind(subset(fd, q==2), e), 
       aes(x=Depth, y=estimate, ymin=LCL, ymax=UCL, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  #stat_smooth(method="lm", fill="gray60")+
  labs(x="Depth (m)", y="Functional Diversity (q = 2)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large

# Linear regression
hl <- splitBy(~Habitat, cbind(subset(fd, q==2), e))
lapply(hl, FUN=function(x)summary(lm(estimate~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = estimate ~ Habitat*Depth, random = ~1|Cruise, data=cbind(subset(fd, q==2), e), method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)

# General linear Models
dat <- cbind(subset(fd, q==2), es)
f <- gls(estimate ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
kable(summary(b)$tTable)
# Relative importance
kable(summary(ma)$sw)
```


```{r, fig.width=6, fig.height=4}
out <- summaryBy(estimate~Cruise+Station+Habitat+Zone, data=cbind(subset(fd, q==2), e), FUN=c(mean, sd))
out$Date <- factor(out$Cruise, labels=c("2015-08", "2015-11"))
names(out)[5:6] <- c("estimate", "sd")

ggplot(data=out, aes(x=Zone, y=estimate, ymin=estimate, ymax=estimate+sd, fill=Date))+
  geom_errorbar(position="dodge")+
  geom_bar(stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Functional Diversity (q = 2)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

# Correlation bettwen functional diversity (q=1, tau=RaoQ) and environmental factor
```{r, fig.width=4.5, fig.height=4}
p1 <- ggplot(data=cbind(subset(fd, q==1), e), 
       aes(x=Speed, y=estimate, ymin=LCL, ymax=UCL))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Current Velocity (m/s)", y="Functional Diversity (q = 1)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~estimate+Speed, data=cbind(subset(fd, q==1), e))
```

```{r, fig.width=4.5, fig.height=4}
p2 <- ggplot(data=cbind(subset(fd, q==1), e), 
       aes(x=over20, y=estimate, ymin=LCL, ymax=UCL))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Duration of Erosion (hr)", y="Functional Diversity (q = 1)")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~estimate+over20, data=cbind(subset(fd, q==1), e))
```

```{r, fig.width=4.5, fig.height=4}
p3 <- ggplot(data=cbind(subset(fd, q==1), e), 
       aes(x=transmissometer, y=estimate, ymin=LCL, ymax=UCL))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Light Transmission (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~estimate+transmissometer, data=cbind(subset(fd, q==1), e))
```


```{r, fig.width=4.5, fig.height=4}
p4 <- ggplot(data=cbind(subset(fd, q==1), e), 
       aes(x=TOC, y=estimate, ymin=LCL, ymax=UCL))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Total Organic Carbon (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~estimate+TOC, data=cbind(subset(fd, q==1), e))
```


```{r, fig.width=10, fig.height=3}
plot_grid(p1, p4, ncol=2, align = "h", axis="b")
```

```{r, fig.width=10, fig.height=3}
plot_grid(p2, p3, ncol=2, align = "h", axis="b")
```

```{r}
out <- cbind(nema_total, "0FD"=subset(fd, q==0)$estimate, "1FD" = subset(fd, q==1)$estimate, "2FD"=subset(fd, q==2)$estimate)
row.names(out) <- NULL
kable(out)
```
