---
title: "Nematode Trophic Diversity and Maturaty Index"
author: "Wei, C.-L., Liao, J.-X."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reguired package
```{r}
library(nema)
library(FD)
library(ggplot2)
library(reshape2)
library(vegan)
library(plyr)
library(viridis)
library(doBy)
library(cowplot)
library(knitr)
library(nlme)
```

# ggplot theme setting
```{r}
large <- theme(legend.title = element_text(size=15),
        legend.text = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text = element_text(size=15),
        strip.text = element_text(size=15))

rotate <- theme(axis.text.x = element_text(angle = 30, hjust=0.5))

no_strip <- theme(strip.background = element_rect(colour=NA, fill=NA),
                  strip.text = element_text(colour=NA))
```

# Preparing nematode and environmental data
Relative abundance of nematode species from 100 randomly slected individuals was scaled to the total nematode abundance in each sample.
```{r}
# Convert to species-by-sample matrix
b <- acast(nema_abund, Cruise+Station+Deployment+Tube+Subcore~Genus+Species, fill=0)
# Multiply the relative abundance by total abundance
b <- ceiling(decostand(b, "total")*nema_total$Abundance)

# Match the trait data to the species-by-sample matrix
s <- nema_species[match(colnames(b), with(nema_species, paste(Genus, Species, sep="_"))),]
```

# Index of trophic diversity & maturity
```{r}
# Buccal morphology
tr <- aggregate(t(b), by=list(s$Buccal), FUN=sum)
buc <- data.frame(t(tr[,-1]))
names(buc) <- tr$Group.1
td <- 1/rowSums(decostand(buc, "total")^2)

# Life history
mi <- colSums(t(decostand(b, "total"))*s$Life.history)

fr <- strsplit(row.names(b), split="_") %>% ldply
out <- with(fr, paste(V1, V2, sep="_")) %>% cbind(fr[,-1:-2], td, mi)
names(out) <- c("Cruise", "Station", "Deployment", "Tube", "Subcore", "td", "mi")
out <- summaryBy(td+mi~Cruise+Station+Deployment, out, FUN=c(mean, sd, length))
names(out)[-1:-3] <- c("td", "mi", "td.sd", "mi.sd", "td.n", "mi.n") 
```

```{r}
# Match environmental data to neamtode sample
id1 <- with(out, paste(Cruise, Station, Deployment))
id2 <- with(nema_cruise, paste(Cruise, Station, Deployment))
out <- cbind(out, nema_cruise[match(id1, id2), -3:-5])

# Define depth zone
depth.bk <- c(200, 400, 600, 800, 1100)
depth.lab <- c("200-400", "400-600", "600-800", "800-1100")
out$Zone <- cut(out$Depth, breaks=depth.bk, labels=depth.lab)
```



```{r, fig.width=6, fig.height=4}
ggplot(data=out, 
       aes(x=Depth, y=td, ymin=td-td.sd, ymax=td+td.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size =5)+
  geom_errorbar()+
  stat_smooth(data=subset(out, Habitat=="Slope"), 
              method="lm", fill="gray60", colour=viridis(2)[2])+
  labs(x="Depth (m)", y="Index of Trophic Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large

# Linear regression
hl <- splitBy(~Habitat, out)
lapply(hl, FUN=function(x)summary(lm(td~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = td ~ Habitat*Depth, random = ~1|Cruise, data=out, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)
```

```{r, fig.width=6, fig.height=4}
out$Date <- factor(out$Cruise, labels=c("2015-08", "2015-11"))

ggplot()+
  geom_errorbar(data=out, aes(x=Zone, y=td, ymin=td, ymax=td+td.sd, fill=Date), position="dodge")+
  geom_bar(data=out, aes(x=Zone, y=td, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Index of Trophic Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```


# Correlation bettwen trophic diversity and environmental factor
```{r, fig.width=4.5, fig.height=4}
p1 <- ggplot(data=out, 
       aes(x=Speed, y=td, ymin=td-td.sd, ymax=td+td.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Current Velocity (m/s)", y="Trophic Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~td+Speed, data=out)
```

```{r, fig.width=4.5, fig.height=4}
p2 <- ggplot(data=out, 
       aes(x=over20, y=td, ymin=td-td.sd, ymax=td+td.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Duration of Erosion (hr)", y="Trophic Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~td+over20, data=out)
```

```{r, fig.width=4.5, fig.height=4}
p3 <- ggplot(data=out, 
       aes(x=transmissometer, y=td, ymin=td-td.sd, ymax=td+td.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Light Transmission (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~td+transmissometer, data=out)
```


```{r, fig.width=4.5, fig.height=4}
p4 <- ggplot(data=out, 
       aes(x=TOC, y=td, ymin=td-td.sd, ymax=td+td.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Total Organic Carbon (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~td+TOC, data=out)
```

```{r, fig.width=10, fig.height=3}
plot_grid(p1, p4, ncol=2, align = "h", axis="b")
```

```{r, fig.width=10, fig.height=3}
plot_grid(p2, p3, ncol=2, align = "h", axis="b")
```

# Maturity index

```{r, fig.width=6, fig.height=4}
ggplot(data=out, 
       aes(x=Depth, y=mi, ymin=mi-mi.sd, ymax=mi+mi.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size =5)+
  geom_errorbar()+
  stat_smooth(data=subset(out, Habitat=="Slope"), 
              method="lm", fill="gray60", colour=viridis(2)[2])+
  labs(x="Depth (m)", y="Maturity Index")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large

# Linear regression
hl <- splitBy(~Habitat, out)
lapply(hl, FUN=function(x)summary(lm(mi~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = mi ~ Habitat*Depth, random = ~1|Cruise, data=out, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)
```

```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=out, aes(x=Zone, y=mi, ymin=mi, ymax=mi+mi.sd, fill=Date), position="dodge")+
  geom_bar(data=out, aes(x=Zone, y=mi, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Maturity Index")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

# Correlation bettwen functional diversity (q=1, tau=RaoQ) and environmental factor
```{r, fig.width=4.5, fig.height=4}
p1 <- ggplot(data=out, 
       aes(x=Speed, y=mi, ymin=mi-mi.sd, ymax=mi+mi.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Current Velocity (m/s)", y="Maturity Index")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~mi+Speed, data=out)
```

```{r, fig.width=4.5, fig.height=4}
p2 <- ggplot(data=out, 
       aes(x=over20, y=mi, ymin=mi-mi.sd, ymax=mi+mi.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Duration of Erosion (hr)", y="Maturity Index")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~mi+over20, data=out)
```

```{r, fig.width=4.5, fig.height=4}
p3 <- ggplot(data=out, 
       aes(x=transmissometer, y=mi, ymin=mi-mi.sd, ymax=mi+mi.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Light Transmission (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~mi+transmissometer, data=out)
```


```{r, fig.width=4.5, fig.height=4}
p4 <- ggplot(data=out, 
       aes(x=TOC, y=mi, ymin=mi-mi.sd, ymax=mi+mi.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Total Organic Carbon (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~mi+TOC, data=out)
```

```{r, fig.width=10, fig.height=3}
plot_grid(p1, p4, ncol=2, align = "h", axis="b")
```

```{r, fig.width=10, fig.height=3}
plot_grid(p2, p3, ncol=2, align = "h", axis="b")
```

```{r}
kable(out[,1:9], align="c")
```
