---
title: "Nematode Diversity standardized by 80.8% sample coverage"
author: "Wei, C.-L., Liao, J.-X."
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

# Reguired package
```{r}
library(nema)
library(iNEXT)
library(ggplot2)
library(reshape2)
library(vegan)
library(plyr)
library(viridis)
library(doBy)
library(cowplot)
library(knitr)
library(nlme)
library(MuMIn)
```

# ggplot theme setting
```{r}
large <- theme(legend.title = element_text(size=15),
        legend.text = element_text(size=15),
        axis.title = element_text(size=15),
        axis.text = element_text(size=15),
        strip.text = element_text(size=15))

rotate <- theme(axis.text.x = element_text(angle = 30, hjust=0.5))

no_strip <- theme(strip.background = element_rect(colour=NA, fill=NA),
                  strip.text = element_text(colour=NA))
```

# Preparing nematode and environmental data
Relative abundance of nematode species from 100 randomly slected individuals was scaled to the total nematode abundance in each sample.
```{r}
# Convert to species-by-sample matrix
b <- acast(nema_abund, Cruise+Station+Deployment+Tube+Subcore~Genus+Species, fill=0)
# Multiply the relative abundance by total abundance
#b <- ceiling(decostand(b, "total")*nema_total$Abundance)

# Match the trait data to the species-by-sample matrix
s <- nema_species[match(colnames(b), with(nema_species, paste(Genus, Species, sep="_"))),]

nema_cruise$Date <- factor(nema_cruise$Cruise, labels=c("2015-08", "2015-11"))
# Define depth zone
depth.bk <- c(200, 400, 600, 800, 1100)
depth.lab <- c("200-400", "400-600", "600-800", "800-1100")
nema_cruise$Zone <- cut(nema_cruise$Depth, breaks=depth.bk, labels=depth.lab)

```
# Abundance
```{r, fig.width=6, fig.height=4}
abu <- summaryBy(Abundance~Cruise+Station+Deployment, nema_total, FUN=c(mean, sd, length))
names(abu)[-1:-3] <- c("Abund", "sd", "n") 
id1 <- with(abu, paste(Cruise, Station, Deployment))
id2 <- with(nema_cruise, paste(Cruise, Station, Deployment))
abu <- cbind(abu, nema_cruise[match(id1, id2), -3:-5])

ggplot(data=abu, 
       aes(x=Depth, y=Abund, ymin=Abund-sd, ymax=Abund+sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5,
                  position=position_jitter(width=10))+
  geom_errorbar()+
  labs(x="Depth (m)", y="Number of Individual")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  scale_y_log10()+
  theme_bw() %+replace% large
```

```{r, fig.width=12, fig.height=12}
# Linear regression
hl <- splitBy(~Habitat, abu)
lapply(hl, FUN=function(x)summary(lm(log10(Abund)~Depth, data=x)))

# Linear Mixed-Effects Model
f <- lme(fixed = log10(Abund) ~ Habitat*Depth, random = ~1|Cruise, data=abu, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)

# Function to plot dianotics plot
dianostic_plot <- 
  function(f, y){
    # standardized residuals versus fitted values
    a1 <- plot(f, resid(., type = "p") ~ fitted(.) | Habitat, abline = 0)
    a2 <- plot(f, resid(., type = "p") ~ fitted(.) | Zone, abline = 0)
    a3 <- plot(f, resid(., type = "p") ~ fitted(.) | Cruise, abline = 0)
    a4 <- plot(f, resid(., type = "p") ~ fitted(.), abline = 0)
    # box-plots of residuals
    b1<-plot(f, Habitat ~ resid(.))
    b2 <-plot(f, Zone ~ resid(.))
    b3 <- plot(f, Cruise ~ resid(.))
    # observed versus fitted values
    c1<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Habitat", sep="|") %>% formula, abline = c(0,1))
    c2<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Zone", sep="|") %>% formula, abline = c(0,1))
    c3<-plot(f, paste(paste(y, "fitted(.)", sep="~"), "Cruise", sep="|") %>% formula, abline = c(0,1))
    c4<-plot(f, paste(y, "fitted(.)", sep="~") %>% formula, abline = c(0,1))
    # QQ plot
    d1<-qqnorm(f, ~ resid(., type = "p") | Habitat, abline = c(0,1))
    d2<-qqnorm(f, ~ resid(., type = "p") | Zone, abline = c(0,1))
    d3<-qqnorm(f, ~ resid(., type = "p") | Cruise, abline = c(0,1))
    d4<-qqnorm(f, ~ resid(., type = "p"), abline = c(0,1))
    
    print(a1, split=c(1,1,4,4), more=TRUE)
    print(a2, split=c(2,1,4,4), more=TRUE)
    print(a3, split=c(3,1,4,4), more=TRUE)
    print(a4, split=c(4,1,4,4), more=TRUE)
    print(b1, split=c(1,2,4,4), more=TRUE)
    print(b2, split=c(2,2,4,4), more=TRUE)
    print(b3, split=c(3,2,4,4), more=TRUE)
    #
    print(c1, split=c(1,3,4,4), more=TRUE)
    print(c2, split=c(2,3,4,4), more=TRUE)
    print(c3, split=c(3,3,4,4), more=TRUE)
    print(c4, split=c(4,3,4,4), more=TRUE)
    print(d1, split=c(1,4,4,4), more=TRUE)
    print(d2, split=c(2,4,4,4), more=TRUE)
    print(d3, split=c(3,4,4,4), more=TRUE)
    print(d4, split=c(4,4,4,4))
    }
dianostic_plot(f, y = "log10(Abund)")


# Adding time into linear model
f <- gls(log10(Abund) ~ Habitat+Depth+Date+Habitat:Depth+Habitat:Date+Depth:Date, data=abu, method = "REML")
summary(f)
dianostic_plot(f, y = "log10(Abund)")

# Pairwise tests on time in canyon
fp <- gls(log10(Abund) ~ Depth+Date+Depth:Date, data=subset(abu, Habitat=="Canyon"), method = "REML")
summary(fp)

# Pairwise tests on time on slope
fp <- gls(log10(Abund) ~ Depth+Date+Depth:Date, data=subset(abu, Habitat=="Slope"), method = "REML")
summary(fp)
```


```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=abu, aes(x=Zone, y=Abund, ymin=Abund, ymax=Abund+sd, fill=Date), position="dodge")+
  geom_bar(data=abu, aes(x=Zone, y=Abund, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Number of Individual")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  scale_y_log10()+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

# Estimate Hill Numbers
* Bootstrap confidence interval (1000 iterations)
```{r, eval=FALSE}
q0 <- iNEXT(t(b), q = 0, size=1:200, nboot=1000)
q1 <- iNEXT(t(b), q = 1, size=1:200, nboot=1000)
q2 <- iNEXT(t(b), q = 2, size=1:200, nboot=1000)
save(list=c("q0", "q1", "q2"), file="../rda/HillNumbers.rda")
```

```{r, fig.width=12, fig.height=4}
load("../rda/HillNumbers.rda")

h <- rbind(ldply(q0$iNextEst, .id="Sample"),
           ldply(q1$iNextEst, .id="Sample"),
           ldply(q2$iNextEst, .id="Sample"))

ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=m, y=qD, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=m, y=qD, group=Sample))+
  geom_vline(xintercept = c(50, 100, 150), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Numbers of Individuals", y = "Effective Number of Species",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip

h <- subset(h, !(order==2 & Sample == "OR1_1126_GC2_1_9_2" | Sample == "OR1_1126_GC2_1_9_3"))

ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=m, y=qD, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=m, y=qD, group=Sample))+
  geom_vline(xintercept = c(50, 100, 150), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Numbers of Individuals", y = "Effective Number of Species",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Size-based diversity accumulation curves based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right). The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_2" and "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r, fig.width=12, fig.height=4}
ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=m, y=SC, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=m, y=SC, group=Sample))+
  geom_vline(xintercept = c(50, 100, 150), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Numbers of Individuals", y = "Sample Coverage",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Sample coverage based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right). The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_2" and "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.


```{r, fig.width=12, fig.height=4}
# Minimum sample coverage for each order q
min(ldply(lapply(splitBy(~order+Sample, subset(h, order==0)), FUN=function(x)x[dim(x)[1],]))$SC)
min(ldply(lapply(splitBy(~order+Sample, subset(h, order==1)), FUN=function(x)x[dim(x)[1],]))$SC)
min(ldply(lapply(splitBy(~order+Sample, subset(h, order==2)), FUN=function(x)x[dim(x)[1],]))$SC)
# SC = 0.807

ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=SC, y=qD, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=SC, y=qD, group=Sample))+
  geom_vline(xintercept = c(0.807), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Sample Coverage", y = "Effective Number of Species",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Coverage-based Diversity accumulation curves based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right) from surface. The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. The vertical dashed line (80.7%) shows the sample with the lowest sample coverage. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_2" and "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r, fig.width=12, fig.height=4}
ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=SC, y=m, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=SC, y=qD, group=Sample))+
  geom_vline(xintercept = c(0.807), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Sample Coverage", y = "Numbers of Individual",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Sample size as a function of sample coverage based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right) from surface. The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. The vertical dashed line (80.7%) shows the sample with the lowest sample coverage. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_2" and "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r, fig.width=12, fig.height=4}
# ~80.7% sample coverage
h0 <- lapply(q0$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  x[1:which(d==min(d)),]
  }) %>% ldply(.id="Sample")

h1 <- lapply(q1$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  x[1:which(d==min(d)),]
  }) %>% ldply(.id="Sample")

h2 <- lapply(q2$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  x[1:which(d==min(d)),]
  }) %>% ldply(.id="Sample")

h <- rbind(h0, h1, h2)
h <- subset(h, !(order==2 & Sample == "OR1_1126_GC2_1_9_3"))

ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=m, y=qD, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=m, y=qD, group=Sample))+
  geom_vline(xintercept = c(50, 100, 150), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Numbers of Individuals", y = "Effective Number of Species",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Size-based diversity accumulation curves based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right). The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. Sample size (numbers of individuals) and the Hill numbers are cut off by 80.7% sample coverage. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r, fig.width=12, fig.height=4}
ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=SC, y=qD, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=SC, y=qD, group=Sample))+
  geom_vline(xintercept = c(0.807), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Sample Coverage", y = "Effective Number of Species",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Coverage-based Diversity accumulation curves based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right) from surface. The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. Sample size (numbers of individuals) and the Hill numbers are cut off by 80.7% sample coverage. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r, fig.width=12, fig.height=4}
ggplot()+
  geom_line(data= subset(h, method!="observed"), 
            aes(x=SC, y=m, colour=method, group=Sample))+
  geom_point(data= subset(h, method=="observed"), 
             aes(x=SC, y=qD, group=Sample))+
  geom_vline(xintercept = c(0.807), linetype=3)+
  facet_wrap(~order, scale="free")+
  labs(x = "Sample Coverage", y = "Numbers of Individual",
       colour="Method")+
  theme_bw()%+replace% large %+replace% no_strip
```

Fig. X. Sample size as a function of sample coverage based on species richness (left), exponential Shannon (middle) and inversed Simpson indices (right) from surface. The blue lines indicate the interpolated (rarefied) and red lines indicate extrapolated parts of the accumulation curves based on 1000 permutations. Dotted symbols indicate the observed diversity values. Sample size (numbers of individuals) and the Hill numbers are cut off by 80.7% sample coverage. The Hill numbers of order q = 2 (inversed Simpson) for sample "OR1_1126_GC2_1_9_3" can not be properly estimated and thus were eliminated.

```{r}
# ~80.7% sample coverage
h0 <- lapply(q0$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  o <- subset(x, d==min(d))
  o[dim(o)[1],]
  }) %>% ldply(.id="Sample")
round(mean(h0$SC), 3)

h1 <- lapply(q1$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  o <- subset(x, d==min(d))
  o[dim(o)[1],]
  }) %>% ldply(.id="Sample")
round(mean(h1$SC), 3)

h2 <- lapply(q2$iNextEst, FUN=function(x) {
  d<-abs(x$SC-0.807)
  o <- subset(x, d==min(d))
  o[dim(o)[1],]
  }) %>% ldply(.id="Sample")
# Set the bad diversity estimates to NA
h2[h2$Sample == "OR1_1126_GC2_1_9_3", c("qD", "qD.LCL", "qD.UCL", "SC", "SC.LCL", "SC.UCL")] <- NA
round(mean(h2$SC, na.rm=TRUE), 3)

# 100 randomly selected individual
#h0 <- subset(ldply(q0$iNextEst, .id="Sample"), m==100)
#h1 <- subset(ldply(q1$iNextEst, .id="Sample"), m==100)
#h2 <- subset(ldply(q2$iNextEst, .id="Sample"), m==100)
# Set the bad diversity estimates to NA
#h2[h2$Sample == "OR1_1126_GC2_1_9_2"|h2$Sample == "OR1_1126_GC2_1_9_3", c("qD", "qD.LCL", "qD.UCL", "SC", "SC.LCL", "SC.UCL")] <- NA
```

## Observed number of species
```{r, fig.width=6, fig.height=4}
fr <- strsplit(as.character(q0$DataInfo$site), split="_") %>% ldply
fr <- cbind(paste(fr$V1, fr$V2, sep="_"), fr[,-1:-2])
names(fr) <- c("Cruise", "Station", "Deployment", "Tube", "Subcore")

div <- cbind(S.obs=q0$DataInfo$S.obs, d0=h0$qD, d1 = h1$qD, d2=h2$qD)
#row.names(div) <- NULL
div <- cbind(fr, div)
div <- summaryBy(S.obs+d0+d1+d2~Cruise+Station+Deployment, div, FUN=c(mean, sd, length))
names(div)[-1:-3] <- c("S.obs", "d0", "d1", "d2", "S.obs.sd", "d0.sd", "d1.sd", "d2.sd", "S.obs.n", "d0.n", "d1.n", "d2.n")

id1 <- with(div, paste(Cruise, Station, Deployment))
id2 <- with(nema_cruise, paste(Cruise, Station, Deployment))
div <- cbind(div, nema_cruise[match(id1, id2), -3:-5])
  
ggplot(data=div, 
       aes(x=Depth, y=S.obs, ymin=S.obs-S.obs.sd, ymax=S.obs+S.obs.sd, colour=Habitat))+
  geom_point(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), size=5)+
  geom_errorbar()+
  stat_smooth(data=subset(div, Habitat=="Slope"), 
       aes(x=Depth, y=S.obs, colour=Habitat), method="lm", fill="gray60")+
  labs(x="Depth (m)", y="Number of Species")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

```{r, fig.width=12, fig.height=12}
# Linear regression
hl <- splitBy(~Habitat, div)
lapply(hl, FUN=function(x)summary(lm(S.obs~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = S.obs ~ Habitat*Depth, random = ~1|Cruise, data=div, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)
dianostic_plot(f, y = "S.obs")

# Adding time into linear model
f <- gls(S.obs ~ Habitat+Depth+Date+Habitat:Depth+Habitat:Date++Depth:Date, data=div, method = "REML")
summary(f)
dianostic_plot(f, y = "S.obs")
# Pairwise tests on time in canyon
fp <- gls(S.obs ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Canyon"), method = "REML")
summary(fp)
# Pairwise tests on time on slope
fp <- gls(S.obs ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Slope"), method = "REML")
summary(fp)

# Normalized enivronmental data to mean zero and unit variance
es <- as.data.frame(scale(div[, c(-1:-23, -45)]))

# General linear Models
dat <- cbind(div[, c(1:23, 45)], es)
f <- gls(S.obs ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
dianostic_plot(f, y = "S.obs")

# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
summary(b)
dianostic_plot(b, y = "S.obs")
kable(summary(b)$tTable)

# Relative importance
kable(summary(ma)$sw)
```


```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=div, aes(x=Zone, y=S.obs, ymin=S.obs, ymax=S.obs+S.obs.sd, fill=Date), position="dodge")+
  geom_bar(data=div, aes(x=Zone, y=S.obs, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Number of Species")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

```{r, fig.width=6, fig.height=4}
div$Abund <- abu$Abund
ggplot(data=div, aes(x=Abund, y=S.obs, colour=Habitat, shape=Date))+
  geom_point(size=5)+
  labs(x="Number of Individual", y="Number of Species")+
  scale_shape_manual(values=c(1, 19))+
  scale_x_log10()+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

## q=0 or species richness
```{r, fig.width=6, fig.height=4}
ggplot(data=div, 
       aes(x=Depth, y=d0, ymin=d0-d0.sd, ymax=d0+d0.sd, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  stat_smooth(method="lm", fill="gray60")+
  labs(x="Depth (m)", y="Species Richness")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

```{r, fig.width=12, fig.height=12}
# Linear regression
hl <- splitBy(~Habitat, div)
lapply(hl, FUN=function(x)summary(lm(d0~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = d0 ~ Habitat*Depth, random = ~1|Cruise, data=div, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)
dianostic_plot(f, y = "d0")

# Adding time into linear model
f <- gls(d0 ~ Habitat+Depth+Date+Habitat:Depth+Habitat:Date++Depth:Date, data=div, method = "REML")
summary(f)
dianostic_plot(f, y = "d0")
# Pairwise tests on time in canyon
fp <- gls(d0 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Canyon"), method = "REML")
summary(fp)
# Pairwise tests on time on slope
fp <- gls(d0 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Slope"), method = "REML")
summary(fp)

# Normalized enivronmental data to mean zero and unit variance
es <- as.data.frame(scale(div[, c(-1:-23, -45)]))

# General linear Models
dat <- cbind(div[, c(1:23, 45:46)], es)
f <- gls(d0 ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
dianostic_plot(f, y = "d0")

# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
summary(b)
dianostic_plot(b, y = "d0")
kable(summary(b)$tTable)

# Relative importance
kable(summary(ma)$sw)
```

```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=div, aes(x=Zone, y=d0, ymin=d0, ymax=d0+d0.sd, fill=Date), position="dodge")+
  geom_bar(data=div, aes(x=Zone, y=d0, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y="Species Richness")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```

```{r, fig.width=6, fig.height=4}
ggplot(data=div, aes(x=Abund, y=d0, colour=Habitat, shape=Date))+
  geom_point(size=5)+
  labs(x="Number of Individual", y="Species Richness")+
  scale_shape_manual(values=c(1, 19))+
  scale_x_log10()+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

## q=1 or exponential Shannon diversity
```{r, fig.width=6, fig.height=4}
ggplot(data=div, 
       aes(x=Depth, y=d1, ymin=d1-d1.sd, ymax=d1+d1.sd, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  stat_smooth(method="lm", fill="gray60")+
  labs(x="Depth (m)", y=expression(exp(Shannon)))+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

```{r, fig.width=12, fig.height=12}
# Linear regression
hl <- splitBy(~Habitat, div)
lapply(hl, FUN=function(x)summary(lm(d0~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = d1 ~ Habitat*Depth, random = ~1|Cruise, data=div, method = "REML", weights=varIdent(form=~1|Cruise))
summary(f)
dianostic_plot(f, y = "d1")

# Adding time into linear model
f <- gls(d1 ~ Habitat+Depth+Date+Habitat:Depth+Habitat:Date++Depth:Date, data=div, method = "REML")
summary(f)
dianostic_plot(f, y = "d1")
# Pairwise tests on time in canyon
fp <- gls(d1 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Canyon"), method = "REML")
summary(fp)
# Pairwise tests on time on slope
fp <- gls(d1 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Slope"), method = "REML")
summary(fp)

# Normalized enivronmental data to mean zero and unit variance
es <- as.data.frame(scale(div[, c(-1:-23, -45)]))

# General linear Models
dat <- cbind(div[, c(1:23, 45:46)], es)
f <- gls(d1 ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
dianostic_plot(f, y = "d1")

# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
summary(b)
dianostic_plot(b, y = "d1")
kable(summary(b)$tTable)

# Relative importance
kable(summary(ma)$sw)
````


```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=div, aes(x=Zone, y=d1, ymin=d1, ymax=d1+d1.sd, fill=Date), position="dodge")+
  geom_bar(data=div, aes(x=Zone, y=d1, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y=expression(exp(Shannon)))+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```


```{r, fig.width=6, fig.height=4}
ggplot(data=div, aes(x=Abund, y=d1, colour=Habitat, shape=Date))+
  geom_point(size=5)+
  labs(x="Number of Individual", y=expression(exp(Shannon)))+
  scale_shape_manual(values=c(1, 19))+
  scale_x_log10()+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

## q=2 or inverse Simpson diversity
```{r, fig.width=6, fig.height=4}
ggplot(data=div, 
       aes(x=Depth, y=d2, ymin=d2-d2.sd, ymax=d2+d2.sd, colour=Habitat))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten = 10, 
                  position=position_jitter(width=10))+
  stat_smooth(method="lm", fill="gray60")+
  labs(x="Depth (m)", y=expression(1/Simpson))+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

```{r, fig.width=12, fig.height=12}
# Linear regression
hl <- splitBy(~Habitat, div)
lapply(hl, FUN=function(x)summary(lm(d0~Depth, data=x)))

# Linear Mixed-Effects Models
f <- lme(fixed = d2 ~ Habitat*Depth, random = ~1|Cruise, data=div, method = "REML", weights=varIdent(form=~1|Cruise), na.action = na.omit)
summary(f)
dianostic_plot(f, y = "d2")

# Adding time into linear model
f <- gls(d2 ~ Habitat+Depth+Date+Habitat:Depth+Habitat:Date++Depth:Date, data=div, method = "REML", na.action = na.omit)
summary(f)
dianostic_plot(f, y = "d2")
# Pairwise tests on time in canyon
fp <- gls(d2 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Canyon"), method = "REML", na.action = na.omit)
summary(fp)
# Pairwise tests on time on slope
fp <- gls(d2 ~ Depth+Date+Depth:Date, data=subset(div, Habitat=="Slope"), method = "REML", na.action = na.omit)
summary(fp)

# Normalized enivronmental data to mean zero and unit variance
es <- as.data.frame(scale(div[, c(-1:-23, -45)]))

# General linear Models
dat <- cbind(div[, c(1:23, 45:46)], es)
dat <- subset(dat, !is.na(d2))

f <- gls(d2 ~ Speed+CN+Salinity+over20+TOC+transmissometer+Temperature, data=dat)
summary(f)
dianostic_plot(f, y = "d2")

# Model selection
ms <- dredge(f)
kable(ms[1:10,])
# Model averaging
ma <- model.avg(ms, fit=TRUE)
kable(summary(ma)$coefmat.full)
# Best model
b <- get.models(ms, subset=1)[[1]]
summary(b)
dianostic_plot(b, y = "d2")
kable(summary(b)$tTable)

# Relative importance
kable(summary(ma)$sw)
````

```{r, fig.width=6, fig.height=4}
ggplot()+
  geom_errorbar(data=div, aes(x=Zone, y=d2, ymin=d2, ymax=d2+d2.sd, fill=Date), position="dodge")+
  geom_bar(data=div, aes(x=Zone, y=d2, fill=Date), stat="identity", position="dodge", colour=gray(0, 0.2))+
  facet_wrap(~Habitat)+
  labs(x="Depth (m)", y=expression(1/Simpson))+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(axis.text.x=element_text(angle=30))
```


```{r, fig.width=6, fig.height=4}
ggplot(data=div, aes(x=Abund, y=d2, colour=Habitat, shape=Date))+
  geom_point(size=5)+
  labs(x="Number of Individual", y=expression(1/Simpson))+
  scale_shape_manual(values=c(1, 19))+
  scale_x_log10()+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large
```

# Correlation bettwen exponential Shannon (q=1) and environmental factor
```{r, fig.width=4.5, fig.height=4}
p1 <- ggplot(data=div, 
       aes(x=Speed, y=d1, ymin=d1-d1.sd, ymax=d1+d1.sd))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Current Velocity (m/s)", y="Shannon Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~d1+Speed, data=div)
```

```{r, fig.width=4.5, fig.height=4}
p2 <- ggplot(data=div, 
       aes(x=over20, y=d1, ymin=d1-d1.sd, ymax=d1+d1.sd))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Modeled Duration of Erosion (hr)", y="Shannon Diversity")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~d1+over20, data=div)
```

```{r, fig.width=4.5, fig.height=4}
p3 <- ggplot(data=div, 
       aes(x=transmissometer, y=d1, ymin=d1-d1.sd, ymax=d1+d1.sd))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Light Transmission (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~d1+transmissometer, data=div)
```

```{r, fig.width=4.5, fig.height=4}
p4 <- ggplot(data=div, 
       aes(x=TOC, y=d1, ymin=d1-d1.sd, ymax=d1+d1.sd))+
  geom_pointrange(aes(fill=Habitat), pch=21, colour=gray(0, 0.2), fatten=10)+
  stat_smooth(method="lm", fill="gray60", colour="gray50")+
  labs(x="Total Organic Carbon (%)", y="")+
  scale_fill_viridis(discrete = TRUE)+
  scale_color_viridis(discrete = TRUE)+
  theme_bw() %+replace% large %+replace% theme(legend.position="none")

cor.test(formula=~d1+TOC, data=div)
```

```{r, fig.width=10, fig.height=3}
plot_grid(p1, p4, ncol=2, align = "h", axis="b")
```

```{r, fig.width=10, fig.height=3}
plot_grid(p2, p3, ncol=2, align = "h", axis="b")
```

```{r}
names(div)[5:7] <- c("0D", "1D", "2D")
names(div)[9:12] <- c("0D.sd", "1D.sd", "2D.sd", "n")
kable(div[,1:12])
```

